Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2023-12-19 paper.tex
Fix a very long formula

Our algorithm fixes $\theta = \frac{pi}{2}$ and does not perform random
rotations or trial triangles. It is sensitive to anisotropy of the medium
when it is stretched along one of the axes as shown in \cref{fig:aniso}. A
direction of elongation for a $N$-dimensional image can be easily detected by
firstly evaluating a $N \times N$ covariance matrix for points belonging to the
phase of interest:
\begin{equation}
  C_{ij} = \sum_{x \in \left\{ \text{Coordinates of points in the
      phase of interest} \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and then decomposing $C$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation and
can be used to align the elongation with the axes (compare \cref{fig:aniso-unrot}
and \cref{fig:aniso-rot}).

The main advantage of our algorithm is its parallelizability and ease of
implementation. An equivalent operation for rotation of test triangles is a
rotation of the whole input image which we have implemented in our
CorrelationFunctions.jl package. By applying a rotation to the input first we
made a rest of the algorithm cache-friendly by avoiding random memory access.
\begin{figure*}[tp]
  \centering
  \subfigure[An original image of an anisotropic pore.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso.png}
    \label{fig:aniso-unrot}}
  \hfill
  \subfigure[A rotated image of pore, so it is elongated along the axis $X$.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso-rot.png}
    \label{fig:aniso-rot}}
  \vskip\baselineskip
  \subfigure[$S_3$ function for the original image]{
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png}}
  \hfill
  \subfigure[$S_3$ function for the rotated image. A vector $(0.92, -0.4)$ is a
    direction of elongation of the pore.]{
  \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png}}
  \hfill
  \subfigure[$S_3$ function averaged across 100 random rotations.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png}
    \label{fig:aniso-avg}}
  \caption[]{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. A result on \cref{fig:aniso-avg} is similar
    to the algorithm proposed by Malmir~et~al.}
  \label{fig:aniso}
\end{figure*}

>>>>

Our algorithm fixes $\theta = \frac{pi}{2}$ and does not perform random
rotations or trial triangles. It is sensitive to anisotropy of the medium
when it is stretched along one of the axes as shown in \cref{fig:aniso}. A
direction of elongation for a $N$-dimensional image can be easily detected by
firstly evaluating a $N \times N$ covariance matrix for points belonging to the
phase of interest:
\begin{equation}
  C_{ij} = \sum_{x \in \left\{
    \begin{array}{l}
      \text{Coordinates of points} \\
      \text{in the phase of interset}
    \end{array}
    \right\}} (x_i - \overline{x_i})(x_j - \overline{x_j})
\end{equation}
and then decomposing $C$ as $T \Lambda T^{-1}$ where $T$ is a unitary matrix and
$\Lambda$ is diagonal. $T$ can always be made so that it defines a rotation and
can be used to align the elongation with the axes (compare \cref{fig:aniso-unrot}
and \cref{fig:aniso-rot}).

The main advantage of our algorithm is its parallelizability and ease of
implementation. An equivalent operation for rotation of test triangles is a
rotation of the whole input image which we have implemented in our
CorrelationFunctions.jl package. By applying a rotation to the input first we
made a rest of the algorithm cache-friendly by avoiding random memory access.
\begin{figure*}[tp]
  \centering
  \subfigure[An original image of an anisotropic pore.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso.png}
    \label{fig:aniso-unrot}}
  \hfill
  \subfigure[A rotated image of the pore, so it is elongated along the axis
    $X$.]{
    \includegraphics[width=0.4\linewidth, frame]{images/aniso-rot.png}
    \label{fig:aniso-rot}}
  \vskip\baselineskip
  \subfigure[$S_3$ function for the original image]{
    \includegraphics[width=0.3\linewidth]{images/aniso-s3.png}}
  \hfill
  \subfigure[$S_3$ function for the rotated image. A vector $(0.92, -0.4)$ is a
    direction of elongation of the pore.]{
  \includegraphics[width=0.3\linewidth]{images/aniso-rot-s3.png}}
  \hfill
  \subfigure[$S_3$ function averaged across 100 random rotations.]{
    \includegraphics[width=0.3\linewidth]{images/aniso-avg-s3.png}
    \label{fig:aniso-avg}}
  \caption[]{Behaviour of our algorithm for calculation of the $S_3$ function
    when dealing with highly anisotropic media. The pattern \cref{fig:pattern}
    which we use in our algorithm is most sensitive to anisotropy which is
    directed along one of the axes. A result on \cref{fig:aniso-avg} is similar
    to the algorithm proposed by Malmir~et~al.}
  \label{fig:aniso}
\end{figure*}
